LUCID: Learning-Enabled Uncertainty-Aware Certification of Stochastic Dynamical Systems

### 2025-08-01 docs/paper/main.tex
feat: move to appendix

\subsection{Data-Driven Dynamics Estimation via Conditional Mean Embeddings}
To reason about the expected value of a random variable, embedding the variable into a (higher dimensional) space and forming a data-driven estimate is a well-established concept in machine learning \cite{scholkopf2002learning,Steinwart2008SVM}.
Following the same reasoning, the \emph{kernel mean embedding} represents the projection of a probability measure into a \gls{rkhs} via a feature map $\phi$ associated with a positive definite kernel $k$~\cite{Smola2007EmbedDistrb,muandet2017kernel}.
For conditional probability distributions a similar concept exists: \glspl{cme} can be used to model the expected value of any \gls{rkhs} function $\new{g}\colon\X\rightarrow\R$ under a stochastic process such as $\S$~\cite{park2020measuretheoretic,muandet2017kernel}.
For the Gaussian kernel,
\begin{equation}
    k(x,x') := \sigma_f^2 \exp\left( -\tfrac{1}{2} (x-x')\T \Sigma\, (x-x') \right),\label{eq:gaussian_kernel}
\end{equation}
where $\Sigma:=\diag(\sigma_l)^{-2}$, with hyperparameters $\sigma_f,\sigma_l\in\R$, \new{the associated RKHS encompasses all smooth functions $g$}.
A data-driven estimate can then be obtained in closed form from a finite amount of data $\data_N$:
\begin{align}
    \begin{split}
         & \E[f(X_+)\mid X=x,\,A=a] \approx                                                  \\
        % \innerH{f}{\hat{\Psi}^N_{X^+|X}(x)}{\Hilbert_{k}} = 
         & \hspace{5em} k_{XA}^N(x,a)\T\left[ K_{XA}^N+N \lambda I_N\right]^{-1}\! f(X_+^N),
    \end{split}\label{eq:empiricalEstimate}
\end{align}
with column vector $k_{XA}^N(x,a):=[k((x^i,a^i),(x,a))]_{i=1}^N$, Gram matrix $K_{XA}^N:=[k((x^i,a^i),(x^j,a^j))]_{i,j=1}^N$, regularization factor $\lambda\geq0$, identity matrix $I_N$, and $f(X_+^N):=[f(x^i_+)]_{i=1}^N$.
The empirical estimator in \eqref{eq:empiricalEstimate} converges in expectation to the true \gls{cme} for $N\rightarrow\infty$~\cite{park2020measuretheoretic}.

@@

\paragraph{Configuration}
% The system and the specification \lucid will check against shall be provided as a configuration to the program.
To certify safety of a system, \lucid accepts a configuration comprising data from the system and the safety specification.
We suggest defining it as a \yaml or equivalent \json file.%
\footnote[4]{The Json schema for \yaml and \json files can be found at \url{https://lucidtoolsource.gitlab.io/lucid/configuration_schema.json}}
% \OS{I would personally avoid the word "scenario", as it could be misunderstood with an actual run of the system.}
If more flexibility is needed, a Python script generating a configuration can be used instead,
with the only requirement being that it must contain a function named \lstinline|scenario_config| returning a \texttt{Configuration} object.
Regardless of their format, configuration files can be loaded with the command \lstinline|pylucid <config file>| to start the tool's verification pipeline.
The same configuration values can also be given directly as command line arguments.
For a list of all available options, run \lstinline|pylucid --help|.
\Sadegh{space needs fixing before ".. to start ...".}

@@

\paragraph{\Circled{4}~Feature Map}
We exploit the truncated spectral kernel expansion in \eqref{eq:sqexp_kernel_fourier} by building an explicit approximated feature map, composed of a linear sum of trigonometric functions with increasing frequencies (see Figure~\ref{fig:spectral_barrier}).
The underlying kernel expansion, visualized in Figure~\ref{fig:spectral_measure_abstraction}, can be controlled to trade-off efficiency and accuracy/conserativeness.
% The full expansion of the kernel would need a sum of infinitely many terms with diminishing impact on the overall result, which we truncate arbitrarily short, a decision based on the tradeoff we are looking for between efficiency and accuracy.
After being defined with the same \hp $(\sigma_f,\sigma_l)$ that characterize the \texttt{KernelRidgeRegressor}, the function can be used to map any point from $\X$ to the \gls{rkhs} associated with the kernel.
\lucid currently implements three different ways for partitioning the spectral measure in Figure~\ref{fig:spectral_measure_abstraction}:
\begin{itemize}
    \item \texttt{LinearTruncatedFourierFeatureMap} defines equally spaced frequency bands on the interval $[-3\sigma_l,3\sigma_l]$, capturing 99.73\% of the spectral measure. This partitioning is shown in Figure~\ref{fig:spectral_measure_abstraction}.
    \item \texttt{LogTruncatedFourierFeatureMap} defines logarithmically spaced frequency bands on the interval $[-3\sigma_l,3\sigma_l]$, with each frequency band carrying an equal share of the spectral measure.
    \item \texttt{ConstantTruncatedFourierFeatureMap} defines frequency bands with a fixed size.
\end{itemize}
Different feature map implementations may yield different results.
In our example, we truncate the Fourier expansion to $5$ frequency bands, including the constant term, and use $64$ times the Nyquist frequency of lattice points to build the constraints of the optimization problem, to avoid it being too conservative.\OS{Specifying it as a multiple of the Nyquist frequency will inevitably raise questions. Just say we generate $X$ lattice points? "lattice size"}
\begin{lstlisting}[language=yaml,backgroundcolor=\color{ipython_bg}]
feature_map: 
    "LinearTruncatedFourierFeatureMap"
num_frequencies: 5
oversample_factor: 64
\end{lstlisting}

@@

\begin{figure}
    \centering
    \resizebox{\columnwidth}{!}{\import{figures}{example.pgf}}
    \caption{Barrier certificate for the running example.
        The state space $\X$ is plotted in the X axis, while the value of the barrier $\B$ is plotted in the Y axis.
        The green line represents the barrier value at time $t$ while the purple one represents the barrier value at time $t+1$.
        The initial and unsafe sets are highlighted in blue and red respectively. \Sadegh{Figure not very readable.}}
    \label{fig:example}
\end{figure}

@@

Since we know the expected behavior of the example system $\S$ from~\eqref{eq:linear-system}, we can run the formal validation by adding the following two lines to the configuration:
\begin{lstlisting}[language=yaml,backgroundcolor=\color{ipython_bg}]
system_dynamics: ["x1 / 2"]
verify: true
\end{lstlisting}
Doing so confirms that $\B$ is indeed a valid \gls{cbc} for the system.

\subsection{Graphical User Interface}

\pylucid provides a \gls{gui} to guide the user in the creation of the scenario configuration and presenting the results in an intuitive way.
Running \lstinline|pylucid-gui| will open a browser tab containing the interface shown in Figure~\ref{fig:gui}, while a local server will listen for requests coming from the \gls{gui}, computing and returning the results.
\todo{Improve picture of the GUI}
\begin{figure}
    \centering
    \includegraphics[trim={0 12.4cm 15cm 0},clip,width=.7\linewidth]{figures/frontend.png}
    \caption{\lucid's \gls{gui} can be accessed via any web browser. 
    The left half of the screen, \Sadegh{as indicated in this picture,} allows the user to provide the transition samples and the safety specification to verify.
    The results will be shown on the right half of the screen.
    }
    \label{fig:gui}
\end{figure}

@@

To ascertain the performance of \lucid, we conduct a series of experiments, run on a Windows 10 machine with an AMD Ryzen 9 5950X 16-Core Processor @ 3.40 GHz, and 64 GB of RAM.
All runs had the random seed set to $42$ to ensure reproducibility.
The results of all experimental results are summarized in Table~\ref{tbl:benchmarks_results}.
\begin{table}[tb]
    \centering
    \begin{tabular}{ccccc}
        \toprule
        \textbf{Benchmark} & \textbf{Dim} & \textbf{Lattice} & \textbf{Runtime} & \textbf{Safety} \\
                           &              & \textbf{Size} & [mm:ss]          & \textbf{Probability}                  \\ % Units
        \midrule
        Linear             & 1            & $?$              & 00:02            & 95.15\%                 \\
        Barr2              & 2            & $350^2$           & 14:42            & 75.34\%                    \\
        Barr3              & 2            & $30^2$           & XX:XX            & XX\%                    \\
                           &              & $40^2$           & XX:XX            & XX\%                    \\
                           &              & $?^2$            & XX:XX            & XX\%                    \\
        Overtaking         & 3            & $?^3$            & XX:XX            & XX\%                    \\
                           & 3            & $?^3$            & XX:XX            & XX\%                    \\
                           & 3            & $?^3$            & XX:XX            & XX\%                    \\
        anotherSimpleOne              & X            &                  & XX:XX            & XX\%                    \\
        \bottomrule
    \end{tabular}
    \caption{Computational benchmarks: ...}
    \label{tbl:benchmarks_results}
\end{table}

\subsection{Barrier 2}

The \barrII experiment is inspired by the one from \cite{abate2021fossil}, which we extend by adding stochastic noise $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_2)$ to arrive at the two-dimensional nonlinear stochastic dynamics

\begin{equation*}
    \begin{bmatrix}
        {x}_{1, t+1} \\
        {x}_{2, t+1}
    \end{bmatrix}
    = \begin{bmatrix}
        {x}_{1, t} \\
        {x}_{2, t}
    \end{bmatrix}+ h\begin{bmatrix}
        {x}_{2, t} - 1 + e^{-x_{1, t}} \\
        -\sin^2(x_{1, t})
    \end{bmatrix} + w_t.
\end{equation*}
with {\color{red}$h:=0.01-0.0001$}.
Given
\begin{align*}
    &\X = [ -2, 2 ] \times [ -2 , 2 ] \\
    &\X_0 = \{ [x_1, x_2] : (x_1 + 0.5)^2 + (x_2 - 0.5)^2 \leq 0.4 \} \\
    &\X_U = \{ [x_1, x_2] : (x_1 - 0.7)^2 + (x_2 + 0.7)^2 \leq 0.3 \}
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=5$ time steps.
\EC{we can remove all the explicit set intervals}
The results are shown in Figure~\ref{fig:CSBarr2}.
The kernel \hp have been selected through offline tuning to be $\sigma_f=1$ and $\sigma_l=[1.0, 5.39334]$.
Using a spectral basis of $M=15^2-1$ frequencies and a lattice density of $\hat{N} = 350^2$, 
we were able to generate a \gls{cbc} with $\eta \approx 5.032$, $\gamma = 100$, and $c \approx 3.926$ in $14.7$ minutes, achieving a lower bound on the safety probability of $P_{\text{safe}}(\S^\pi) \geq 75.34\%$.
% Knowing the system dynamics, we were also able to formally validate the barrier.

\begin{figure}[ht]
    %\resizebox{\columnwidth}{!}{\import{figures}{barrier2.pdf}}%
    \includegraphics[trim={5cm 3cm 2cm 4cm},clip,width=\linewidth]{barrier2.pdf}
    \caption{\gls{cbc} synthesis for the \barrII benchmark.
        The blue circle represents the initial set $\X_0$, while the red one represents the unsafe set $\X_U$.
    }
    \label{fig:CSBarr2}
\end{figure}

\subsection{Barrier 3}

Similarly, we also tested \lucid on the \barrIII benchmark adapted from \cite{abate2021fossil} with added stochastic noise.
The resulting nonlinear stochastic dynamics are
\begin{equation*}
    \begin{bmatrix}
        {x}_{1, t+1} \\
        {x}_{2, t+1}
    \end{bmatrix}
    = \begin{bmatrix}
        {x}_{1, t} \\
        {x}_{2, t}
    \end{bmatrix}+ h
    \begin{bmatrix}
        {x}_{2, t} \\
        \frac{1}{3} {x}^3_{1, t} - {x}_{1,t} - {x}_{2,t}
    \end{bmatrix} + w_t
\end{equation*}
with {\color{red}$h:=0.01-0.0001$}.
Given
\begin{align*}
    &\X = [ -3, 2.5 ] \times [ -2 , 1 ] \\
    &\X_0 = [ 1 , 2 ] \times [ -0.7 , 0.3 ] \cup [ -1.8 , -1.4 ] \times [ -0.1 , 0.1 ] \\
    &\qquad \cup [-1.4, -1.2] \times [-0.5 , 0.1] \\
    &\X_U = [ 0.4 , 0.6 ] \times [ 0.2 , 0.6 ] \cup [ 0.6 , 0.7 ] \times [ 0.2 , 0.4 ],
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=10$ time steps.
\EC{we can remove all the explicit set intevals}
This example is particularly challenging due to the closeness between the initial set and the unsafe set.
Hence, we need a rather high number of frequencies to achieve the required expressivity in the barrier.
The results are shown in Figure~\ref{fig:CSBarr3}.
The kernel \hp have been selected through offline tuning to be $\sigma_f=X$ and $\sigma_l=[X, X]$.
Using a spectral basis of $M=X^3-1$ frequencies and a lattice density of $\hat{N} = X^2$, 
we were able to generate a \gls{cbc} with $\eta \approx X.X$, $\gamma \approx X.X$, and $c \approx X.X$ in $X$ minutes and $X$ seconds, achieving a safety probability of $P_{safe}(\S^\pi) \approx XX.X\%$.
% Knowing the system dynamics, we were also able to formally validate the barrier.

\begin{figure}[ht]
    \resizebox{\columnwidth}{!}{\import{figures}{barrier3.pgf}}%
    \caption{\gls{cbc} synthesis for the \barrIII benchmark.
        The blue regions represent the initial set $\X_0$, while the red regions represent the unsafe set $\X_U$.
        The pink surface indicates the value of $\gamma$.
    }
    \label{fig:CSBarr3}
\end{figure}

\subsection{Overtaking}

We consider a scenario where an autonomous vehicle (AV) controlled by a \gls{nn} is overtaking another vehicle, as indicated in Figure~\ref{fig:safety_setup}. 
The dynamics of the ego vehicle are given by Dubin's car model \EC{does it need a cit or is it just textbook} by adding the noise vector $w = \begin{bmatrix}w_t^1 & w_t^2 & w_t^3\end{bmatrix}$ where each component is drawn from a zero-mean Gaussian with standard deviation $0.01$, $0.01$, and $0.001$, respectively.
The steering wheel angle is supplied by the \gls{nn} controller and we travel at a fixed velocity $v=1$.
The goal is to certify that the controlled system remains safe by computing the probability that the system initialized in $[{x}_0,{y}_0,\phi_0]\T\in \X_0$ does not enter the unsafe region $\X_U$ within $T=50$ time steps.

The \gls{lp} is generated based on a lattice of $\hat{N}=X^3$ equidistant points and a spectral basis of $M=4^3-1$ frequencies.
The kernel \hp are selected to be $\sigma_f=X$ and $\sigma_l=[X, X, X]$.
We obtained a barrier which certifies safety with a probability of at least $XX\%$ with $c=0$, $\gamma=X.X$ and $\eta=X.X$.
Note that this is only certifying safety w.r.t. the empirical distribution, i.e., the \gls{cme} constructed from the observed data.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{placeholder.jpeg}
    \caption{Barrier}
    \label{fig:barr3_barrier}
\end{figure}

@@

\section{Conclusion and Future Extensions}\label{sec:conclusion_and_future_extensions}
\Sadegh{we should combine Conclusion and Future extensions into one section and keep it in the main body. We should also mention sparse CME as a way of improving scalability...}

\new{This paper introduces \lucid, a novel }

The form of data \eqref{eq:data} considered here assumes access to full state measurements, as it is the case when working with simulated systems. Extending this to partially observed settings adds another level of complexity. Extending \lucid to such settings is on our agenda and builds on top of the results presented in this paper.

\OS{As the GaussianKernel and KernelRidgeRegressor are the only options we currently provide, is specifying them explicitly necessary? Maybe we could also add a sentence about the fact that they are currently the only option, but that in the future we plan to extend this to to allow to include prior knowledge about the system dynamics (in the spirit of physics-informed approaches).}

\OS{Mention robustness radius}

\section{Conclusion}\label{sec:conclusion}
...

>>>>

\subsection{Data-Driven Dynamics Estimation via Conditional Mean Embeddings}\label{sec:dd_estimation_via_cme}
To reason about the expected value of a random variable, embedding the variable into a (higher dimensional) space and forming a data-driven estimate is a well-established concept in machine learning \cite{scholkopf2002learning,Steinwart2008SVM}.
Following the same reasoning, the \emph{kernel mean embedding} represents the projection of a probability measure into a \gls{rkhs} via a feature map $\phi$ associated with a positive definite kernel $k$~\cite{Smola2007EmbedDistrb,muandet2017kernel}.
For conditional probability distributions a similar concept exists: \glspl{cme} can be used to model the expected value of any \gls{rkhs} function $\new{g}\colon\X\rightarrow\R$ under a stochastic process such as $\S$~\cite{park2020measuretheoretic,muandet2017kernel}.
For the Gaussian kernel,
\begin{equation}
    k(x,x') := \sigma_f^2 \exp\left( -\tfrac{1}{2} (x-x')\T \Sigma\, (x-x') \right),\label{eq:gaussian_kernel}
\end{equation}
where $\Sigma:=\diag(\sigma_l)^{-2}$, with hyperparameters $\sigma_f,\sigma_l\in\R$, \new{the associated RKHS encompasses all smooth functions $g$}.
A data-driven estimate can then be obtained in closed form from a finite amount of data $\data_N$:
\begin{align}
    \begin{split}
         & \E[f(X_+)\mid X=x,\,A=a] \approx                                                  \\
        % \innerH{f}{\hat{\Psi}^N_{X^+|X}(x)}{\Hilbert_{k}} = 
         & \hspace{5em} k_{XA}^N(x,a)\T\left[ K_{XA}^N+N \lambda I_N\right]^{-1}\! f(X_+^N),
    \end{split}\label{eq:empiricalEstimate}
\end{align}
with column vector $k_{XA}^N(x,a):=[k((x^i,a^i),(x,a))]_{i=1}^N$, Gram matrix $K_{XA}^N:=[k((x^i,a^i),(x^j,a^j))]_{i,j=1}^N$, regularization factor $\lambda\geq0$, identity matrix $I_N$, and $f(X_+^N):=[f(x^i_+)]_{i=1}^N$.
The empirical estimator in \eqref{eq:empiricalEstimate} converges in expectation to the true \gls{cme} for $N\rightarrow\infty$~\cite{park2020measuretheoretic}.

@@

\paragraph{Configuration}
% The system and the specification \lucid will check against shall be provided as a configuration to the program.
To certify safety of a system, \lucid accepts a configuration comprising data from the system and the safety specification.
We suggest defining it as a \yaml or equivalent \json file.%
\footnote[4]{The Json schema for \yaml and \json files can be found at \url{https://lucidtoolsource.gitlab.io/lucid/configuration_schema.json}}
% \OS{I would personally avoid the word "scenario", as it could be misunderstood with an actual run of the system.}
If more flexibility is needed, a Python script generating a configuration can be used instead,
with the only requirement being that it must contain a function named \lstinline|scenario_config| returning a \texttt{Configuration} object.
Regardless of their format, configuration files can be loaded with the command \lstinline|pylucid <config file>| to start the tool's verification pipeline.
The same configuration values can also be given directly as command line arguments.
\pylucid also provides a browser-based \gls{gui} to aid in the configuration and execution of the tool.
\Sadegh{space needs fixing before ".. to start ...".}

@@

\paragraph{\Circled{4}~Feature Map}
We exploit the truncated spectral kernel expansion in \eqref{eq:sqexp_kernel_fourier} by building an explicit approximated feature map, composed of a linear sum of trigonometric functions with increasing frequencies (see Figure~\ref{fig:spectral_barrier}).
The underlying kernel expansion, visualized in Figure~\ref{fig:spectral_measure_abstraction}, can be controlled to trade-off efficiency and accuracy/conserativeness.
% The full expansion of the kernel would need a sum of infinitely many terms with diminishing impact on the overall result, which we truncate arbitrarily short, a decision based on the tradeoff we are looking for between efficiency and accuracy.
After being defined with the same \hp $(\sigma_f,\sigma_l)$ that characterize the \texttt{KernelRidgeRegressor}, the function can be used to map any point from $\X$ to the \gls{rkhs} associated with the kernel.
\lucid currently implements three different ways for partitioning the spectral measure in Figure~\ref{fig:spectral_measure_abstraction}:
\begin{itemize}
    \item \texttt{LinearTruncatedFourierFeatureMap} defines equally spaced frequency bands on the interval $[-3\sigma_l,3\sigma_l]$, capturing 99.73\% of the spectral measure. This partitioning is shown in Figure~\ref{fig:spectral_measure_abstraction}.
    \item \texttt{LogTruncatedFourierFeatureMap} defines logarithmically spaced frequency bands on the interval $[-3\sigma_l,3\sigma_l]$, with each frequency band carrying an equal share of the spectral measure.
    \item \texttt{ConstantTruncatedFourierFeatureMap} defines frequency bands with a fixed size.
\end{itemize}
Different feature map implementations may yield different results.
In our example, we truncate the Fourier expansion to $5$ frequency bands, including the constant term, and use $704$ lattice points to build the constraints of the optimization problem, to avoid it being too conservative.\OS{Specifying it as a multiple of the Nyquist frequency will inevitably raise questions. Just say we generate $X$ lattice points? "lattice size"}\EC{new}
\begin{lstlisting}[language=yaml,backgroundcolor=\color{ipython_bg}]
feature_map: 
    "LinearTruncatedFourierFeatureMap"
num_frequencies: 5
num_oversample: 704
\end{lstlisting}

@@

\begin{figure}
    \centering
    \begin{adjustbox}{max width=\columnwidth}
        \input{figures/example.pgf}
    \end{adjustbox}
    %\resizebox{\columnwidth}{!}{\import{figures}{example.pgf}}
    \caption{Barrier certificate for the running example.
        The state space $\X$ is plotted in the X axis, while the value of the barrier $\B$ is plotted in the Y axis.
        The green line represents the barrier value at time $t$.
        The initial and unsafe sets are highlighted in blue and red respectively. % \Sadegh{Figure not very readable.}
        }
    \label{fig:example}
\end{figure}

@@

Since we know the expected behavior of the example system $\S$ from~\eqref{eq:linear-system}, we can run the formal validation by adding the following two lines to the configuration:
\begin{lstlisting}[language=yaml,backgroundcolor=\color{ipython_bg}]
system_dynamics: ["x1 / 2"]
verify: true
\end{lstlisting}
Doing so confirms that $\B$ is indeed a valid \gls{cbc}.

@@

To ascertain the performance of \lucid, we conduct a series of experiments, run on a Windows 10 machine with an AMD Ryzen 9 5950X 16-Core Processor @ 3.40 GHz, NVIDIA GeForce RTX 3090 GPU, and 64 GB of RAM.
All runs had the random seed set to $42$ to ensure reproducibility.

We adapt the \barrII and \barrIII benchmarks from \cite{abate2021fossil}.
Both are two-dimensional highly nonlinear systems to which we add the stochastic noise $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_2)$.
Given the initial set $\X_0$ and the unsafe set $\X_U$, we synthesize a barrier $\B$ that guarantees trajectories starting in $\X_0$ do not enter $\X_U$ within $T=5$ time steps.
Note that we can only certify safety w.r.t. the empirical distribution, i.e., the \gls{cme} constructed from the observed data.
\barrIII is particularly challenging due to the closeness between the initial set and the unsafe set.
The synthesized barriers are shown in Figures~\ref{fig:CSBarr2}--\ref{fig:CSBarr3}.
We also consider a scenario, we name \overtaking, where an autonomous vehicle (AV) controlled by a \gls{nn} is overtaking another vehicle.
The dynamics of the ego vehicle are given by Dubin's car model with an added noise vector $w$ where each component is drawn from a zero-mean Gaussian with standard deviation $0.01$, $0.01$, and $0.001$ respectively.
The steering wheel angle is supplied by the \gls{nn} controller and we travel at a fixed velocity.

Table~\ref{tbl:benchmarks_results} summarizes the results of all the experiments presented.
More details about the configurations and \hp used can be found in the Appendix.

\begin{figure}[ht]
    \begin{adjustbox}{clip,trim=0cm 1.7cm 0cm 2cm,max width=\columnwidth}
        \input{figures/barrier2.pgf}
    \end{adjustbox}
    %\includegraphics[trim={5cm 3cm 2cm 4cm},clip,width=\linewidth]{barrier2.pgf}
    \caption{\gls{cbc} synthesised for the \barrII benchmark.}
    \label{fig:CSBarr2}
\end{figure}

\begin{figure}[ht]
    \begin{adjustbox}{clip,trim=0cm 2cm 0cm 2cm,max width=\columnwidth}
        \input{figures/barrier3.pgf}
    \end{adjustbox}
    %\resizebox{\columnwidth}{!}{\import{figures}{barrier3.pgf}}%
    \caption{\gls{cbc} synthesised for the \barrIII benchmark.}
    \label{fig:CSBarr3}
\end{figure}

\begin{table}[tb]
    \centering
    \begin{tabular}{cccccc}
        \toprule
                    & \textbf{T} & \textbf{Freq} & \textbf{Lattice} & \textbf{Runtime} & \textbf{Safety} \\
                    &            &               & \textbf{Size}    & [mm:ss]          & \textbf{Prob}   \\ % Units
        \midrule
        \linear     & 15         & 5             & $704$            & 00:02            & 93.74\%         \\
        \barrII     & 5          & 15            & $350^2$          & 10:01            & 56.87\%         \\
        \barrIII    & 5          & 15            & $800^2$          & 64:20            & 42.25\%         \\
        \overtaking & 5          & 5             & $99^3$           & 46:33            & 30.32\%         \\
        \bottomrule
    \end{tabular}
    \caption{Computational benchmarks.
        The lattice size indicates the number of points of the lattice as the resolution for each dimension raised to the number of dimensions.
        $T$ is the number of time steps associated with the lower bound on the safety probability, displayed in the last column.
    }
    \label{tbl:benchmarks_results}
\end{table}

% Linear
% Success: 93.74%, c 0.0032813476604243406, eta 0.013421457997340467, lambda 1e-05, num_frequencies 5, num_oversample 704, oversample_factor 2.0, sigma_l 0.034, sigma_f 18.0, T 15
% Barrier3
% Success: 42.25%, c 0.11928419647787443, eta 0.5584972599863635, lambda 1e-08, num_frequencies 15, num_oversample 800, oversample_factor 1.0, sigma_l [2.99266 4.62946], sigma_f 1.0, T 5
% Overtaking
% Success: 30.32%, c 0.05031756638044985, eta 0.09681458793205037, lambda 1e-05, num_frequencies 5, num_oversample -1, oversample_factor 9.0, sigma_l [10.  7.  5.], sigma_f 7.0, T 5

@@

@@

\section{Conclusion and Future Extensions}\label{sec:conclusion_and_future_extensions}
\new{This paper introduces \lucid, a novel tool capable of quantifying safety guarantees for black-box systems with complex stochastic dynamics and deep learning components in the loop. As such, \lucid fills a gap currently unaddressed by preexisting tools. To achieve this, \lucid, leverages Conditional Mean Embeddings (CME) to learn control barrier certificates from data, and recasts the problem into a tractable linear form via a spectral abstraction.}

\new{
    \lucid is built for extensibility. In its current form, it assumes access to full state measurements (cf.~\eqref{eq:data}), as is the case when working with simulated systems, and focuses on verification, i.e., when a policy is given.
    Extending \lucid to partially observed settings and safe controller synthesis
    % adds another level of complexity and 
    is on our agenda and builds on top of the results presented in this paper.
    \lucid derives barriers based on the empirical CME, which can be robustified against out-of-sample system behavior by tightening the barrier constraint (c) (cf.~Sections~\ref{sec:dd_estimation_via_cme}--\ref{sec:ddbarriers}).
    The scalability of \lucid can be improved by performing sparse CME computations or specializing the estimator's \texttt{Kernel} to embed prior knowledge about the system dynamics.
}
